---
title: "MP4"
author: "Sarah Gillespie, Berry Williams, Eva Putman"
date: "2019-05-04"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r StartingItems, message=FALSE, warning = FALSE}
library(mdsr)
library(RMySQL)
library(tidyverse)
library(dplyr)
```

```{r SQLData, message=FALSE, warning = FALSE}
library(RMySQL)
db <- dbConnect(MySQL(), 
                host = "scidb.smith.edu", 
                user = "mth292", 
                password = "RememberPi", 
                dbname = "citibike")

class(db)

```
### Notes
### Mimic the Wonder Woman article
# i really like the station_summary that shows the most popular stations
# we could count the number of rides total and by station.

#trip_summary just shows an increasing number of people are using the bikes. Heartening but not particularly insightful.
#try for a single bike then see where it goes
# somebody else did analysis on this but using Python: https://towardsdatascience.com/citi-bike-2017-analysis-efd298e6c22c
```{r ActiveCode, message=FALSE, warning = FALSE}
db %>%
  dbGetQuery("SHOW TABLES;")

db %>%
  dbGetQuery("SELECT * 
             FROM station_months
             LIMIT 0, 100;")

db %>%
  dbGetQuery("SELECT * 
             FROM station_summary
             LIMIT 0, 100;")
db %>%
  dbGetQuery("SELECT * 
             FROM trip_summary
             LIMIT 0, 20;")

db %>%
  dbGetQuery("SELECT * 
              FROM trips
              LIMIT 0, 200;")

```

```{r ActiveCode, message=FALSE, warning = FALSE}

```

#### Questions: let's go at this from a business perspective. Are most bike riders customers or subscribers, age, gender. How can we get more people to ride the bikes? What geographic location do the bikes start and end?

```{r ActiveCode, message=FALSE, warning = FALSE}

joinedTable <- db %>%
  dbGetQuery("SELECT
                station_months.station_id, station_months.name, station_months.num_starts, station_months.num_stops, station_summary.num_months, station_summary.num_starts, station_summary.num_stops
              FROM
                station_months
                  LEFT JOIN
                station_summary ON station_months.station_id = station_summary.station_id;")
```

^[[Github Repository](https://github.com/SarahGillespie/sds192-mp4)
]
